Scientific Reasoning with LLMs

### 2024-10-29 docs/proposal/main.tex
RQ2
final version

\documentclass[10pt]{article}
\usepackage[a4paper, margin=1.25in]{geometry}

@@

\paragraph{Prompting techniques}. A prompt is a command a user gives to a generative LLM used to guide its output for a downstream task, which contains several of the following: a role (persona) the LLM has to follow (e.g. "You are a helpful assistant"), a precise task description, examples (exemplars) for analogous learning (few-shot prompting), requirements on the process of reasoning, style instructions, output format requirements, additional information for in-context learning, emotion prompting components (highlighting why the task or a particular requirement is important). Prompts can be combined into sequences or graphs (with complex branching and parallelism). Of particular interest are small requirements on the reasoning process that can significantly improve performance when added to the prompt, such as 
%"Rephrase and expand the question, and respond" \cite{}, or 
"Let’s think step by step" \cite{kojima2022large}. 
Chain-of-thought prompts improve reasoning capabilities by asking the model to speak its process of thinking out loud, with variations including self-ask, step-back prompting, thread-of-thought, least-to-most prompting (decomposing then solving), plan-and-solve prompting, tree-of-thought, recursion-of-thought and many more \cite{schulhoff2024prompt}. 
%with the resulting systems possibly using multiple CoT reasoning trajectories to select the most likely one (e.g. Uncertainty-Routed CoT Prompting, or making an LLM generate various prompting instructions that are then tested by another LLM on 24 NLP tasks to chose the best one \cite{zhou2022large}). %katja: I omit this to save space as not directly relevant
Furthermore, prompts themselves can be compressed and automatically optimized to improve efficiency and reduce costs \cite{chang2024efficient}. \cite{ge2023context} compress a long context 4x into memory slots, while \cite{weston2023system} ask the LLM to first summarize the prompt and then execute it. Self-criticism and ensembling techniques can further be used to improve reasoning capabilities \cite{schulhoff2024prompt}.

@@

\paragraph{Agents}. An important step forward in using LLMs are agent-based architectures \cite{lin2024swiftsage} \cite{ghafarollahi2024sciagents}. They allow the LLM to act as an intelligent being capable of planning a solution of a complex task, while resorting to external resources on the way to acquire in-context the knowledge it does not have from pre-training. The retrieved information, results of invoking tools, such as calculators, and of interactions with the environment are appended to the prompt (or, in a more advanced scenario, to the summarized memory). Optionally, a reflexion step (explicit reasoning on all the accumulated information) is added before allowing further generation of the final response.

@@

Scientific reasoning and the dataset \cite{lu2022learn}

\section{Methodology}
We postulate the following research questions: 
RQ1: Can we build a LLM Agent to improve the LLMs performance on science questions?
RQ2: Does Soft Prompting improve the performance of the LLM (Agent)?
RQ3: Can an LLM-agent's behavior be distilled into a single model?
\paragraph{RQ1} 
We plan to

@@

\paragraph{Metrics}
We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards. A few metrics used in machine translation such as BLEU-1, BLEU-4 and BERTScore can also be used, but the main metric we'll rely on is accuracy for question answering.

\section{Approximate Timeline}
-Now - mid November: 
  Preliminary baseline tests: small LLM relying only on commonsense reasoning / only on pretraining 
  zero-shot, few-shot
  optional: stronger baseline – same LLM with better prompting techniques like CoT/Self-Ask
-mid November - mid December:
  Construct agentic / RAG pipeline with “big smart” model + additional resources.
  optional: soft-prompting for better workflow^ to distill from
-mid December - mid January:
  Knowledge distillation experiments.

\katja{Have reverted the doc to simple "article"}

>>>>

\documentclass[10pt]{article}
\usepackage[a4paper, margin=1.0in]{geometry}

@@

\paragraph{Prompting techniques} A prompt is a command a user gives to a generative LLM used to guide its output for a downstream task, which contains several of the following: a role (persona) the LLM has to follow (e.g. "You are a helpful assistant"), a precise task description, examples (exemplars) for analogous learning (few-shot prompting), requirements on the process of reasoning, style instructions, output format requirements, additional information for in-context learning, emotion prompting components (highlighting why the task or a particular requirement is important). Prompts can be combined into sequences or graphs (with complex branching and parallelism). Of particular interest are small requirements on the reasoning process that can significantly improve performance when added to the prompt, such as 
%"Rephrase and expand the question, and respond" \cite{}, or 
"Let’s think step by step" \cite{kojima2022large}. 
Chain-of-thought prompts improve reasoning capabilities by asking the model to speak its process of thinking out loud, with variations including self-ask, step-back prompting, thread-of-thought, least-to-most prompting (decomposing then solving), plan-and-solve prompting, tree-of-thought, recursion-of-thought and many more \cite{schulhoff2024prompt}. 
%with the resulting systems possibly using multiple CoT reasoning trajectories to select the most likely one (e.g. Uncertainty-Routed CoT Prompting, or making an LLM generate various prompting instructions that are then tested by another LLM on 24 NLP tasks to chose the best one \cite{zhou2022large}). %katja: I omit this to save space as not directly relevant
Furthermore, prompts themselves can be compressed and automatically optimized to improve efficiency and reduce costs \cite{chang2024efficient}. \cite{ge2023context} compress a long context 4x into memory slots, while \cite{weston2023system} ask the LLM to first summarize the prompt and then execute it. Self-criticism and ensembling techniques can further be used to improve reasoning capabilities \cite{schulhoff2024prompt}.

@@

\paragraph{Agents} An important step forward in using LLMs are agent-based architectures \cite{lin2024swiftsage} \cite{ghafarollahi2024sciagents}. They allow the LLM to act as an intelligent being capable of planning a solution of a complex task, while resorting to external resources on the way to acquire in-context the knowledge it does not have from pre-training. The retrieved information, results of invoking tools, such as calculators, and of interactions with the environment are appended to the prompt (or, in a more advanced scenario, to the summarized memory). Optionally, a reflexion step (explicit reasoning on all the accumulated information) is added before allowing further generation of the final response.

@@

\section{Methodology}
We postulate the following research questions: 
RQ1: Can we build a LLM Agent to improve the LLMs performance on science questions?
RQ2: Does Soft Prompting improve the performance of the LLM (Agent)?
RQ3: Can an LLM-agent's behavior be distilled into a single model?

\paragraph{RQ1} 
We plan to start by prompting a multimodal model with reasoning capabilities with a zero-shot and few-shot prompting settings. A foundation model for this step could be a T4/T5 model or a multimodal Llama model. This simple generation will be compared with an agent-based approach using the same model, which will now include augmented retrieval of information on the scientific task from domain-specific texts.

@@

\paragraph{Metrics}
We'll evaluate model's performance with question answering accuracy domain-wise in order to have a fair comparison with leaderboards. A few metrics used in machine translation such as BLEU-1, BLEU-4 and BERTScore can also be used to evaluate the reasoning steps.
%, but the main metric we'll rely on is accuracy for question answering.

\section{Approximate Timeline}
\begin{enumerate}
	\item \textbf{Now - mid November}: 
		\begin{itemize}
  			\item Preliminary baseline tests: small LLM relying only on commonsense reasoning / only on pretraining (zero-shot, few-shot).
			\item Optional: stronger baseline – same LLM with better prompting techniques like CoT/Self-Ask.
		\end{itemize}
	\item \textbf{mid November - mid December}:
		\begin{itemize}
  			\item Construct agentic / RAG pipeline with “big smart” model + additional resources.
  			\item Optional: soft-prompting for better workflow to distill from.
		\end{itemize}
	\item \textbf{mid December - mid January}:
  		\begin{itemize}
			\item Knowledge distillation experiments.
		\end{itemize}
\end{enumerate}
